---
title: "SuSiE-RSS Fine-Mapping for Dental Traits"
author: "Divya Pandey"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---

## Overview

This analysis applies SuSiE-RSS fine-mapping to dental GWAS summary statistics and LD matrices.

The current workflow is implemented for **DMFS and Nteeth** and includes:

### DMFS

1. SuSiE credible set results  
2. Positional mapping to nearest hg19 genes  
3. eQTL colocalization summary  

### Nteeth

1. SuSiE credible set results  
2. Positional mapping to nearest hg19 genes  
3. eQTL colocalization summary  

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(DT)
```

# DMFS

## DMFS SuSiE credible set results

```{r dmfs-susie-results, echo=TRUE}
dmfs_susie_final_summary <- read.csv("results/dmfs_susie_final_summary.csv")

dmfs_susie_final_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    MarkerName,
    CS,
    PIP,
    Purity,
    everything()
  ) %>%
  datatable(
    extensions = "Buttons",
    caption = "DMFS SuSiE Credible Set Results",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## DMFS position mapping

```{r dmfs-position-mapping, echo=TRUE}
dmfs_positional_mapping <- read.csv("results/dmfs_positional_mapping.csv")

dmfs_positional_mapping %>%
  dplyr::select(
    Locus,
    MarkerName,
    PIP,
    external_gene_name,
    ensembl_gene_id,
    everything()
  ) %>%
  arrange(desc(PIP)) %>%
  datatable(
    caption = "DMFS SuSiE Results with Nearest hg19 Genes",
    extensions = "Buttons",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## DMFS eQTL colocalization

```{r dmfs-eqtl-results, echo=TRUE}
dmfs_eqtl_summary <- readLines("results/dmfs_eqtl_colocalization_summary.txt")

cat(dmfs_eqtl_summary)
```

# Nteeth

## Nteeth SuSiE credible set results

```{r nteeth-susie-results, echo=TRUE}
nteeth_susie_final_summary <- read.csv("results/nteeth_susie_final_summary.csv")

nteeth_susie_final_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    MarkerName,
    CS,
    PIP,
    Purity,
    everything()
  ) %>%
  datatable(
    extensions = "Buttons",
    caption = "Nteeth SuSiE Credible Set Results",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## Nteeth position mapping

```{r nteeth-position-mapping, echo=TRUE}
nteeth_positional_mapping <- read.csv("results/nteeth_positional_mapping.csv")

nteeth_positional_mapping %>%
  dplyr::select(
    Locus,
    MarkerName,
    PIP,
    external_gene_name,
    ensembl_gene_id,
    everything()
  ) %>%
  arrange(desc(PIP)) %>%
  datatable(
    caption = "Nteeth SuSiE Results with Nearest hg19 Genes",
    extensions = "Buttons",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## Nteeth eQTL colocalization

```{r nteeth-eqtl-results, echo=TRUE}
nteeth_eqtl_summary <- readLines("results/nteeth_eqtl_colocalization_summary.txt")

cat(nteeth_eqtl_summary)
```
# Joint mvSuSiE Fine-Mapping: DMFS and Nteeth

This section presents the joint multi-trait fine-mapping results for DMFS and Nteeth using mvSuSiE-RSS. GWAS Z-scores for both traits were aligned to the same locus-specific EUR LD matrices. LD SNP identifiers were mapped from rsID to chr:position format using the corresponding `.id_map` files.

```{r joint-mvsusie-results, echo=TRUE, message=FALSE, warning=FALSE}
mvsusie_summary <- read.csv("results/dmfs_nteeth_mvsusie_summary.csv")

mvsusie_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    rsid,
    MarkerName,
    CS,
    PIP,
    Purity,
    Z_DMFS,
    Z_Nteeth,
    Pvalue_DMFS,
    Pvalue_Nteeth,
    Effect_DMFS,
    Effect_Nteeth,
    everything()
  ) %>%
  datatable(
    extensions = "Buttons",
    caption = "Joint mvSuSiE Fine-Mapping Results for DMFS and Nteeth",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## Top joint mvSuSiE signals

```{r top-joint-mvsusie-results, echo=TRUE}
mvsusie_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    rsid,
    MarkerName,
    CS,
    PIP,
    Purity,
    Z_DMFS,
    Z_Nteeth,
    Pvalue_DMFS,
    Pvalue_Nteeth
  ) %>%
  head(20)
```