---
title: "SuSiE-RSS Fine-Mapping for Dental Traits"
author: "Divya Pandey"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---

## Overview

This analysis applies SuSiE-RSS fine-mapping to dental GWAS summary statistics and LD matrices.

The current workflow is implemented for **DMFS and Nteeth** and includes:

### DMFS

1. SuSiE credible set results  
2. Positional mapping to nearest hg19 genes  
3. eQTL colocalization summary  

### Nteeth

1. SuSiE credible set results  
2. Positional mapping to nearest hg19 genes  
3. eQTL colocalization summary  

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(DT)
```

## SuSiE credible set results

```{r susie-results, echo=TRUE}
susie_final_summary <- read.csv("results/dmfs_susie_final_summary.csv")

susie_final_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    MarkerName,
    CS,
    PIP,
    Purity,
    everything()
  ) %>%
  datatable(
    extensions = "Buttons",
    caption = "DMFS SuSiE Credible Set Results",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## Position mapping

```{r position-mapping, echo=TRUE}
positional_mapping <- read.csv("results/dmfs_positional_mapping.csv")

positional_mapping %>%
  dplyr::select(
    Locus,
    MarkerName,
    PIP,
    external_gene_name,
    ensembl_gene_id,
    everything()
  ) %>%
  arrange(desc(PIP)) %>%
  datatable(
    caption = "SuSiE Results with Nearest hg19 Genes",
    extensions = "Buttons",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## eQTL colocalization

```{r eqtl-results, echo=TRUE}
eqtl_summary <- readLines("results/dmfs_eqtl_colocalization_summary.txt")

cat(eqtl_summary)
```