---
title: "SuSiE-RSS Fine-Mapping for Dental Traits"
author: "Divya Pandey"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---

## Overview

This analysis applies SuSiE-RSS and mvSuSiE-RSS fine-mapping to dental GWAS summary statistics and locus-specific LD matrices.

The GWAS summary statistics used in this project come from Shungin et al., **“Genome-wide analysis of dental caries and periodontitis combining clinical and self-reported data,”** published in *Nature Communications* in 2019. This study analyzed dental caries and periodontal traits using clinical measures from the GLIDE consortium and self-reported oral health traits from UK Biobank. The clinical dental traits used in this fine-mapping workflow include **DMFS**, **DFSS**, and **Nteeth**.

Article: [Shungin et al. 2019, Nature Communications](https://www.nature.com/articles/s41467-019-10630-1)

The current workflow is implemented for **DMFS, Nteeth, and DFSS** and includes:

### Single-trait SuSiE-RSS

1. SuSiE credible set results  
2. Positional mapping to nearest hg19 genes  
3. eQTL colocalization summary  

### Joint mvSuSiE-RSS

1. DMFS + Nteeth  
2. DMFS + DFSS  
3. Nteeth + DFSS  
4. DMFS + Nteeth + DFSS  

For joint mvSuSiE-RSS analyses, only variants with **PIP = 1** are displayed, with nearest hg19 gene annotation added from the positional mapping results.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(DT)

make_dt <- function(data, caption_text) {
  datatable(
    data,
    extensions = "Buttons",
    caption = caption_text,
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
}

make_nearest_genes <- function(...) {
  bind_rows(...) %>%
    dplyr::select(
      MarkerName,
      external_gene_name,
      ensembl_gene_id,
      any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene"))
    ) %>%
    distinct(MarkerName, .keep_all = TRUE)
}
```

# DMFS

## DMFS SuSiE credible set results

```{r dmfs-susie-results, echo=TRUE}
dmfs_susie_final_summary <- read.csv(
  "results/dmfs_susie_final_summary.csv"
)

dmfs_susie_final_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    MarkerName,
    CS,
    PIP,
    Purity,
    everything()
  ) %>%
  make_dt("DMFS SuSiE Credible Set Results")
```

## DMFS position mapping

```{r dmfs-position-mapping, echo=TRUE}
dmfs_positional_mapping <- read.csv(
  "results/dmfs_positional_mapping.csv"
)

dmfs_positional_mapping %>%
  dplyr::select(
    Locus,
    MarkerName,
    PIP,
    external_gene_name,
    ensembl_gene_id,
    everything()
  ) %>%
  arrange(desc(PIP)) %>%
  make_dt("DMFS SuSiE Results with Nearest hg19 Genes")
```

## DMFS eQTL colocalization

```{r dmfs-eqtl-results, echo=TRUE}
dmfs_eqtl_summary <- readLines(
  "results/dmfs_eqtl_colocalization_summary.txt"
)

cat(dmfs_eqtl_summary)
```

# Nteeth

## Nteeth SuSiE credible set results

```{r nteeth-susie-results, echo=TRUE}
nteeth_susie_final_summary <- read.csv(
  "results/nteeth_susie_final_summary.csv"
)

nteeth_susie_final_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    MarkerName,
    CS,
    PIP,
    Purity,
    everything()
  ) %>%
  make_dt("Nteeth SuSiE Credible Set Results")
```

## Nteeth position mapping

```{r nteeth-position-mapping, echo=TRUE}
nteeth_positional_mapping <- read.csv(
  "results/nteeth_positional_mapping.csv"
)

nteeth_positional_mapping %>%
  dplyr::select(
    Locus,
    MarkerName,
    PIP,
    external_gene_name,
    ensembl_gene_id,
    everything()
  ) %>%
  arrange(desc(PIP)) %>%
  make_dt("Nteeth SuSiE Results with Nearest hg19 Genes")
```

## Nteeth eQTL colocalization

```{r nteeth-eqtl-results, echo=TRUE}
nteeth_eqtl_summary <- readLines(
  "results/nteeth_eqtl_colocalization_summary.txt"
)

cat(nteeth_eqtl_summary)
```

# DFSS

## DFSS SuSiE credible set results

```{r dfss-susie-results, echo=TRUE}
dfss_susie_final_summary <- read.csv(
  "results/dfss_susie_final_summary.csv"
)

dfss_susie_final_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    MarkerName,
    CS,
    PIP,
    Purity,
    everything()
  ) %>%
  make_dt("DFSS SuSiE Credible Set Results")
```

## DFSS position mapping

```{r dfss-position-mapping, echo=TRUE}
dfss_positional_mapping <- read.csv(
  "results/dfss_positional_mapping.csv"
)

dfss_positional_mapping %>%
  dplyr::select(
    Locus,
    MarkerName,
    PIP,
    external_gene_name,
    ensembl_gene_id,
    everything()
  ) %>%
  arrange(desc(PIP)) %>%
  make_dt("DFSS SuSiE Results with Nearest hg19 Genes")
```

## DFSS eQTL colocalization

```{r dfss-eqtl-results, echo=TRUE}
cat(
  "No eQTL matches found in the specified tissue: Whole_Blood"
)
```

# Joint mvSuSiE-RSS Fine-Mapping

## Joint analysis workflow

For the joint mvSuSiE-RSS analyses, the same general workflow was used across all trait pairs and the three-trait analysis:

1. GWAS summary statistics were aligned to the same locus-specific EUR LD matrices.
2. The corresponding `.id_map` files were used to map LD rsIDs to chromosome-position marker names.
3. Z-scores were calculated as `Effect / StdErr`.
4. mvSuSiE-RSS was run using the same prior variance structure and `L = 10`.
5. Joint PIP, credible set assignment, purity, Z-scores, effect estimates, standard errors, p-values, sample sizes, and nearest hg19 gene annotation were summarized.

For all joint mvSuSiE-RSS analyses, only variants with `PIP = 1` are displayed in the tables below.

Model convergence was checked from the saved mvSuSiE-RSS fit objects. Some loci did not converge even after increasing the maximum number of iterations to 1000, so results from non-converged loci should be interpreted cautiously. PIP values were verified to remain bounded between 0 and 1 for all joint summary outputs.

The mvSuSiE-RSS scripts are available here:


- [DMFS + Nteeth: `07_mvsusie_dmfs_nteeth.R`](https://github.com/divpandey715/dental-genetics-susie/blob/main/scripts/07_mvsusie_dmfs_nteeth.R)
- [DMFS + DFSS: `11_mvsusie_dmfs_dfss.R`](https://github.com/divpandey715/dental-genetics-susie/blob/main/scripts/11_mvsusie_dmfs_dfss.R)
- [Nteeth + DFSS: `12_mvsusie_nteeth_dfss.R`](https://github.com/divpandey715/dental-genetics-susie/blob/main/scripts/12_mvsusie_nteeth_dfss.R)
- [DMFS + Nteeth + DFSS: `13_mvsusie_dmfs_nteeth_dfss.R`](https://github.com/divpandey715/dental-genetics-susie/blob/main/scripts/13_mvsusie_dmfs_nteeth_dfss.R)

## Joint mvSuSiE Fine-Mapping: DMFS and Nteeth

This section presents the joint multi-trait fine-mapping results for DMFS and Nteeth using mvSuSiE-RSS.

```{r joint-mvsusie-dmfs-nteeth, echo=TRUE, message=FALSE, warning=FALSE}
nearest_genes_dmfs_nteeth <- make_nearest_genes(
  dmfs_positional_mapping,
  nteeth_positional_mapping
)

dmfs_nteeth_trait_cols <- c(
  "Z_DMFS",
  "Z_Nteeth",
  "Pvalue_DMFS",
  "Pvalue_Nteeth",
  "Effect_DMFS",
  "Effect_Nteeth",
  "StdErr_DMFS",
  "StdErr_Nteeth",
  "N_DMFS",
  "N_Nteeth"
)

dmfs_nteeth_mvsusie_summary <- read.csv(
  "results/dmfs_nteeth_mvsusie_summary.csv"
)

dmfs_nteeth_mvsusie_summary_pip1 <- dmfs_nteeth_mvsusie_summary %>%
  filter(PIP == 1) %>%
  dplyr::select(
    -any_of(c(
      "external_gene_name",
      "ensembl_gene_id",
      "distance_to_gene",
      "distance",
      "distance_to_nearest_gene"
    ))
  ) %>%
  left_join(nearest_genes_dmfs_nteeth, by = "MarkerName") %>%
  distinct() %>%
  dplyr::select(
    any_of(c("Locus", "rsid", "MarkerName", "CS", "PIP", "Purity")),
    external_gene_name,
    ensembl_gene_id,
    any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene")),
    any_of(dmfs_nteeth_trait_cols),
    everything()
  ) %>%
  arrange(Locus, MarkerName)

dmfs_nteeth_mvsusie_summary_pip1 %>%
  make_dt("Joint mvSuSiE Fine-Mapping Results for DMFS and Nteeth: PIP = 1 with Nearest hg19 Genes")
```

### Top joint DMFS and Nteeth mvSuSiE signals

```{r top-joint-mvsusie-dmfs-nteeth, echo=TRUE}
dmfs_nteeth_mvsusie_summary_pip1 %>%
  dplyr::select(
    any_of(c(
      "Locus",
      "rsid",
      "MarkerName",
      "CS",
      "PIP",
      "Purity",
      "external_gene_name",
      "ensembl_gene_id",
      "Z_DMFS",
      "Z_Nteeth",
      "Pvalue_DMFS",
      "Pvalue_Nteeth"
    ))
  ) %>%
  head(20)
```

## Joint mvSuSiE Fine-Mapping: DMFS and DFSS

This section presents the joint multi-trait fine-mapping results for DMFS and DFSS using the same mvSuSiE-RSS workflow.

```{r joint-mvsusie-dmfs-dfss, echo=TRUE, message=FALSE, warning=FALSE}
nearest_genes_dmfs_dfss <- make_nearest_genes(
  dmfs_positional_mapping,
  dfss_positional_mapping
)

dmfs_dfss_trait_cols <- c(
  "Z_DMFS",
  "Z_DFSS",
  "Pvalue_DMFS",
  "Pvalue_DFSS",
  "Effect_DMFS",
  "Effect_DFSS",
  "StdErr_DMFS",
  "StdErr_DFSS",
  "N_DMFS",
  "N_DFSS"
)

dmfs_dfss_mvsusie_summary <- read.csv(
  "results/dmfs_dfss_mvsusie_summary.csv"
)

dmfs_dfss_mvsusie_summary_pip1 <- dmfs_dfss_mvsusie_summary %>%
  filter(PIP == 1) %>%
  dplyr::select(
    -any_of(c(
      "external_gene_name",
      "ensembl_gene_id",
      "distance_to_gene",
      "distance",
      "distance_to_nearest_gene"
    ))
  ) %>%
  left_join(nearest_genes_dmfs_dfss, by = "MarkerName") %>%
  distinct() %>%
  dplyr::select(
    any_of(c("Locus", "rsid", "MarkerName", "CS", "PIP", "Purity")),
    external_gene_name,
    ensembl_gene_id,
    any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene")),
    any_of(dmfs_dfss_trait_cols),
    everything()
  ) %>%
  arrange(Locus, MarkerName)

dmfs_dfss_mvsusie_summary_pip1 %>%
  make_dt("Joint mvSuSiE Fine-Mapping Results for DMFS and DFSS: PIP = 1 with Nearest hg19 Genes")
```

### Top joint DMFS and DFSS mvSuSiE signals

```{r top-joint-mvsusie-dmfs-dfss, echo=TRUE}
dmfs_dfss_mvsusie_summary_pip1 %>%
  dplyr::select(
    any_of(c(
      "Locus",
      "rsid",
      "MarkerName",
      "CS",
      "PIP",
      "Purity",
      "external_gene_name",
      "ensembl_gene_id",
      "Z_DMFS",
      "Z_DFSS",
      "Pvalue_DMFS",
      "Pvalue_DFSS"
    ))
  ) %>%
  head(20)
```

## Joint mvSuSiE Fine-Mapping: Nteeth and DFSS

This section presents the joint multi-trait fine-mapping results for Nteeth and DFSS using the same mvSuSiE-RSS workflow.

```{r joint-mvsusie-nteeth-dfss, echo=TRUE, message=FALSE, warning=FALSE}
nearest_genes_nteeth_dfss <- make_nearest_genes(
  nteeth_positional_mapping,
  dfss_positional_mapping
)

nteeth_dfss_trait_cols <- c(
  "Z_Nteeth",
  "Z_DFSS",
  "Pvalue_Nteeth",
  "Pvalue_DFSS",
  "Effect_Nteeth",
  "Effect_DFSS",
  "StdErr_Nteeth",
  "StdErr_DFSS",
  "N_Nteeth",
  "N_DFSS"
)

nteeth_dfss_mvsusie_summary <- read.csv(
  "results/nteeth_dfss_mvsusie_summary.csv"
)

nteeth_dfss_mvsusie_summary_pip1 <- nteeth_dfss_mvsusie_summary %>%
  filter(PIP == 1) %>%
  dplyr::select(
    -any_of(c(
      "external_gene_name",
      "ensembl_gene_id",
      "distance_to_gene",
      "distance",
      "distance_to_nearest_gene"
    ))
  ) %>%
  left_join(nearest_genes_nteeth_dfss, by = "MarkerName") %>%
  distinct() %>%
  dplyr::select(
    any_of(c("Locus", "rsid", "MarkerName", "CS", "PIP", "Purity")),
    external_gene_name,
    ensembl_gene_id,
    any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene")),
    any_of(nteeth_dfss_trait_cols),
    everything()
  ) %>%
  arrange(Locus, MarkerName)

nteeth_dfss_mvsusie_summary_pip1 %>%
  make_dt("Joint mvSuSiE Fine-Mapping Results for Nteeth and DFSS: PIP = 1 with Nearest hg19 Genes")
```

### Top joint Nteeth and DFSS mvSuSiE signals

```{r top-joint-mvsusie-nteeth-dfss, echo=TRUE}
nteeth_dfss_mvsusie_summary_pip1 %>%
  dplyr::select(
    any_of(c(
      "Locus",
      "rsid",
      "MarkerName",
      "CS",
      "PIP",
      "Purity",
      "external_gene_name",
      "ensembl_gene_id",
      "Z_Nteeth",
      "Z_DFSS",
      "Pvalue_Nteeth",
      "Pvalue_DFSS"
    ))
  ) %>%
  head(20)
```

## Joint mvSuSiE Fine-Mapping: DMFS, Nteeth, and DFSS

This section presents the three-trait joint mvSuSiE-RSS fine-mapping results for DMFS, Nteeth, and DFSS.

```{r joint-mvsusie-all-three, echo=TRUE, message=FALSE, warning=FALSE}
nearest_genes_all_three <- make_nearest_genes(
  dmfs_positional_mapping,
  nteeth_positional_mapping,
  dfss_positional_mapping
)

all_three_trait_cols <- c(
  "Z_DMFS",
  "Z_Nteeth",
  "Z_DFSS",
  "Pvalue_DMFS",
  "Pvalue_Nteeth",
  "Pvalue_DFSS",
  "Effect_DMFS",
  "Effect_Nteeth",
  "Effect_DFSS",
  "StdErr_DMFS",
  "StdErr_Nteeth",
  "StdErr_DFSS",
  "N_DMFS",
  "N_Nteeth",
  "N_DFSS"
)

dmfs_nteeth_dfss_mvsusie_summary <- read.csv(
  "results/dmfs_nteeth_dfss_mvsusie_summary.csv"
)

dmfs_nteeth_dfss_mvsusie_summary_pip1 <- dmfs_nteeth_dfss_mvsusie_summary %>%
  filter(PIP == 1) %>%
  dplyr::select(
    -any_of(c(
      "external_gene_name",
      "ensembl_gene_id",
      "distance_to_gene",
      "distance",
      "distance_to_nearest_gene"
    ))
  ) %>%
  left_join(nearest_genes_all_three, by = "MarkerName") %>%
  distinct() %>%
  dplyr::select(
    any_of(c("Locus", "rsid", "MarkerName", "CS", "PIP", "Purity")),
    external_gene_name,
    ensembl_gene_id,
    any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene")),
    any_of(all_three_trait_cols),
    everything()
  ) %>%
  arrange(Locus, MarkerName)

dmfs_nteeth_dfss_mvsusie_summary_pip1 %>%
  make_dt("Joint mvSuSiE Fine-Mapping Results for DMFS, Nteeth, and DFSS: PIP = 1 with Nearest hg19 Genes")
```

### Top joint DMFS, Nteeth, and DFSS mvSuSiE signals

```{r top-joint-mvsusie-all-three, echo=TRUE}
dmfs_nteeth_dfss_mvsusie_summary_pip1 %>%
  dplyr::select(
    any_of(c(
      "Locus",
      "rsid",
      "MarkerName",
      "CS",
      "PIP",
      "Purity",
      "external_gene_name",
      "ensembl_gene_id",
      "Z_DMFS",
      "Z_Nteeth",
      "Z_DFSS",
      "Pvalue_DMFS",
      "Pvalue_Nteeth",
      "Pvalue_DFSS"
    ))
  ) %>%
  head(20)
```
