---
title: "SuSiE-RSS Fine-Mapping for Dental Traits"
author: "Divya Pandey"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---

## Overview

This analysis applies SuSiE-RSS fine-mapping to dental GWAS summary statistics and LD matrices.

The current workflow is implemented for **DMFS, Nteeth, and DFSS** and includes:

### DMFS

1. SuSiE credible set results  
2. Positional mapping to nearest hg19 genes  
3. eQTL colocalization summary  

### Nteeth

1. SuSiE credible set results  
2. Positional mapping to nearest hg19 genes  
3. eQTL colocalization summary  

### DFSS

1. SuSiE credible set results  
2. Positional mapping to nearest hg19 genes  
3. eQTL colocalization summary  

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(DT)
```

# DMFS

## DMFS SuSiE credible set results

```{r dmfs-susie-results, echo=TRUE}
dmfs_susie_final_summary <- read.csv(
  "results/dmfs_susie_final_summary.csv"
)

dmfs_susie_final_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    MarkerName,
    CS,
    PIP,
    Purity,
    everything()
  ) %>%
  datatable(
    extensions = "Buttons",
    caption = "DMFS SuSiE Credible Set Results",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## DMFS position mapping

```{r dmfs-position-mapping, echo=TRUE}
dmfs_positional_mapping <- read.csv(
  "results/dmfs_positional_mapping.csv"
)

dmfs_positional_mapping %>%
  dplyr::select(
    Locus,
    MarkerName,
    PIP,
    external_gene_name,
    ensembl_gene_id,
    everything()
  ) %>%
  arrange(desc(PIP)) %>%
  datatable(
    caption = "DMFS SuSiE Results with Nearest hg19 Genes",
    extensions = "Buttons",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## DMFS eQTL colocalization

```{r dmfs-eqtl-results, echo=TRUE}
dmfs_eqtl_summary <- readLines(
  "results/dmfs_eqtl_colocalization_summary.txt"
)

cat(dmfs_eqtl_summary)
```

# Nteeth

## Nteeth SuSiE credible set results

```{r nteeth-susie-results, echo=TRUE}
nteeth_susie_final_summary <- read.csv(
  "results/nteeth_susie_final_summary.csv"
)

nteeth_susie_final_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    MarkerName,
    CS,
    PIP,
    Purity,
    everything()
  ) %>%
  datatable(
    extensions = "Buttons",
    caption = "Nteeth SuSiE Credible Set Results",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## Nteeth position mapping

```{r nteeth-position-mapping, echo=TRUE}
nteeth_positional_mapping <- read.csv(
  "results/nteeth_positional_mapping.csv"
)

nteeth_positional_mapping %>%
  dplyr::select(
    Locus,
    MarkerName,
    PIP,
    external_gene_name,
    ensembl_gene_id,
    everything()
  ) %>%
  arrange(desc(PIP)) %>%
  datatable(
    caption = "Nteeth SuSiE Results with Nearest hg19 Genes",
    extensions = "Buttons",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## Nteeth eQTL colocalization

```{r nteeth-eqtl-results, echo=TRUE}
nteeth_eqtl_summary <- readLines(
  "results/nteeth_eqtl_colocalization_summary.txt"
)

cat(nteeth_eqtl_summary)
```

# DFSS

## DFSS SuSiE credible set results

```{r dfss-susie-results, echo=TRUE}
dfss_susie_final_summary <- read.csv(
  "results/dfss_susie_final_summary.csv"
)

dfss_susie_final_summary %>%
  arrange(desc(PIP)) %>%
  dplyr::select(
    Locus,
    MarkerName,
    CS,
    PIP,
    Purity,
    everything()
  ) %>%
  datatable(
    extensions = "Buttons",
    caption = "DFSS SuSiE Credible Set Results",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## DFSS position mapping

```{r dfss-position-mapping, echo=TRUE}
dfss_positional_mapping <- read.csv(
  "results/dfss_positional_mapping.csv"
)

dfss_positional_mapping %>%
  dplyr::select(
    Locus,
    MarkerName,
    PIP,
    external_gene_name,
    ensembl_gene_id,
    everything()
  ) %>%
  arrange(desc(PIP)) %>%
  datatable(
    caption = "DFSS SuSiE Results with Nearest hg19 Genes",
    extensions = "Buttons",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## DFSS eQTL colocalization

```{r dfss-eqtl-results, echo=TRUE}
cat(
  "No eQTL matches found in the specified tissue: Whole_Blood"
)
```

# Joint mvSuSiE Fine-Mapping: DMFS and Nteeth

This section presents the joint multi-trait fine-mapping results for DMFS and Nteeth using mvSuSiE-RSS. Only variants with **PIP = 1** are displayed, with nearest hg19 gene annotation added from the positional mapping results.

```{r joint-mvsusie-results, echo=TRUE, message=FALSE, warning=FALSE}
mvsusie_summary <- read.csv(
  "results/dmfs_nteeth_mvsusie_summary.csv"
)

dmfs_genes <- read.csv("results/dmfs_positional_mapping.csv")
nteeth_genes <- read.csv("results/nteeth_positional_mapping.csv")

nearest_genes_dmfs_nteeth <- bind_rows(
  dmfs_genes,
  nteeth_genes
) %>%
  dplyr::select(
    MarkerName,
    external_gene_name,
    ensembl_gene_id,
    any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene"))
  ) %>%
  distinct(MarkerName, .keep_all = TRUE)

mvsusie_summary_pip1 <- mvsusie_summary %>%
  filter(PIP == 1) %>%
  left_join(nearest_genes_dmfs_nteeth, by = "MarkerName") %>%
  distinct() %>%
  dplyr::select(
    Locus,
    rsid,
    MarkerName,
    CS,
    PIP,
    Purity,
    external_gene_name,
    ensembl_gene_id,
    any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene")),
    Z_DMFS,
    Z_Nteeth,
    Pvalue_DMFS,
    Pvalue_Nteeth,
    Effect_DMFS,
    Effect_Nteeth,
    everything()
  ) %>%
  arrange(Locus, MarkerName)

mvsusie_summary_pip1 %>%
  datatable(
    extensions = "Buttons",
    caption = "Joint mvSuSiE Fine-Mapping Results for DMFS and Nteeth: PIP = 1 with Nearest hg19 Genes",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## Top joint mvSuSiE signals

```{r top-joint-mvsusie-results, echo=TRUE}
mvsusie_summary_pip1 %>%
  dplyr::select(
    Locus,
    rsid,
    MarkerName,
    CS,
    PIP,
    Purity,
    external_gene_name,
    ensembl_gene_id,
    Z_DMFS,
    Z_Nteeth,
    Pvalue_DMFS,
    Pvalue_Nteeth
  ) %>%
  head(20)
```

# Joint mvSuSiE Fine-Mapping: DMFS and DFSS

This section presents the joint multi-trait fine-mapping results for DMFS and DFSS using mvSuSiE-RSS. Only variants with **PIP = 1** are displayed, with nearest hg19 gene annotation added from the positional mapping results.

```{r joint-mvsusie-dmfs-dfss, echo=TRUE, message=FALSE, warning=FALSE}
dmfs_dfss_mvsusie_summary <- read.csv(
  "results/dmfs_dfss_mvsusie_summary.csv"
)

dmfs_genes <- read.csv("results/dmfs_positional_mapping.csv")
dfss_genes <- read.csv("results/dfss_positional_mapping.csv")

nearest_genes_dmfs_dfss <- bind_rows(
  dmfs_genes,
  dfss_genes
) %>%
  dplyr::select(
    MarkerName,
    external_gene_name,
    ensembl_gene_id,
    any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene"))
  ) %>%
  distinct(MarkerName, .keep_all = TRUE)

dmfs_dfss_mvsusie_summary_pip1 <- dmfs_dfss_mvsusie_summary %>%
  filter(PIP == 1) %>%
  left_join(nearest_genes_dmfs_dfss, by = "MarkerName") %>%
  distinct() %>%
  dplyr::select(
    Locus,
    rsid,
    MarkerName,
    CS,
    PIP,
    Purity,
    external_gene_name,
    ensembl_gene_id,
    any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene")),
    Z_DMFS,
    Z_DFSS,
    Pvalue_DMFS,
    Pvalue_DFSS,
    Effect_DMFS,
    Effect_DFSS,
    everything()
  ) %>%
  arrange(Locus, MarkerName)

dmfs_dfss_mvsusie_summary_pip1 %>%
  datatable(
    extensions = "Buttons",
    caption = "Joint mvSuSiE Fine-Mapping Results for DMFS and DFSS: PIP = 1 with Nearest hg19 Genes",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## Top joint DMFS and DFSS mvSuSiE signals

```{r top-joint-mvsusie-dmfs-dfss, echo=TRUE}
dmfs_dfss_mvsusie_summary_pip1 %>%
  dplyr::select(
    Locus,
    rsid,
    MarkerName,
    CS,
    PIP,
    Purity,
    external_gene_name,
    ensembl_gene_id,
    Z_DMFS,
    Z_DFSS,
    Pvalue_DMFS,
    Pvalue_DFSS
  ) %>%
  head(20)
```

# Joint mvSuSiE Fine-Mapping: Nteeth and DFSS

This section presents the joint multi-trait fine-mapping results for Nteeth and DFSS using mvSuSiE-RSS. Only variants with **PIP = 1** are displayed, with nearest hg19 gene annotation added from the positional mapping results.

```{r joint-mvsusie-nteeth-dfss, echo=TRUE, message=FALSE, warning=FALSE}
nteeth_dfss_mvsusie_summary <- read.csv(
  "results/nteeth_dfss_mvsusie_summary.csv"
)

nteeth_genes <- read.csv("results/nteeth_positional_mapping.csv")
dfss_genes <- read.csv("results/dfss_positional_mapping.csv")

nearest_genes_nteeth_dfss <- bind_rows(
  nteeth_genes,
  dfss_genes
) %>%
  dplyr::select(
    MarkerName,
    external_gene_name,
    ensembl_gene_id,
    any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene"))
  ) %>%
  distinct(MarkerName, .keep_all = TRUE)

nteeth_dfss_mvsusie_summary_pip1 <- nteeth_dfss_mvsusie_summary %>%
  filter(PIP == 1) %>%
  left_join(nearest_genes_nteeth_dfss, by = "MarkerName") %>%
  distinct() %>%
  dplyr::select(
    Locus,
    rsid,
    MarkerName,
    CS,
    PIP,
    Purity,
    external_gene_name,
    ensembl_gene_id,
    any_of(c("distance_to_gene", "distance", "distance_to_nearest_gene")),
    Z_Nteeth,
    Z_DFSS,
    Pvalue_Nteeth,
    Pvalue_DFSS,
    Effect_Nteeth,
    Effect_DFSS,
    everything()
  ) %>%
  arrange(Locus, MarkerName)

nteeth_dfss_mvsusie_summary_pip1 %>%
  datatable(
    extensions = "Buttons",
    caption = "Joint mvSuSiE Fine-Mapping Results for Nteeth and DFSS: PIP = 1 with Nearest hg19 Genes",
    options = list(
      dom = "Blfrtip",
      buttons = c("copy", "csv", "excel", "pdf", "print"),
      lengthMenu = list(
        c(10, 25, 50, -1),
        c(10, 25, 50, "All")
      ),
      scrollX = TRUE
    )
  )
```

## Top joint Nteeth and DFSS mvSuSiE signals

```{r top-joint-mvsusie-nteeth-dfss, echo=TRUE}
nteeth_dfss_mvsusie_summary_pip1 %>%
  dplyr::select(
    Locus,
    rsid,
    MarkerName,
    CS,
    PIP,
    Purity,
    external_gene_name,
    ensembl_gene_id,
    Z_Nteeth,
    Z_DFSS,
    Pvalue_Nteeth,
    Pvalue_DFSS
  ) %>%
  head(20)
```
